---
title: "Question 3.4 for Presentation"
author: "Kai Lukowiak"
date: '2017-09-19'
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 3.4 Triathlon times, Part I.
In triathlons, it is common for racers to be placed into age
and gender groups. Friends Leo and Mary both completed the Hermosa Beach Triathlon, where
Leo competed in the Men, Ages 30 - 34 group while Mary competed in the Women, Ages 25 -
29 group. Leo completed the race in 1:22:28 (4948 seconds), while Mary completed the race in
1:31:53 (5513 seconds). Obviously Leo finished faster, but they are curious about how they did
within their respective groups. Can you help them? Here is some information on the performance
of their groups:

* The finishing times of the Men, Ages 30 - 34 group has a mean of 4313 seconds with a
standard deviation of 583 seconds.

* The finishing times of the Women, Ages 25 - 29 group has a mean of 5261 seconds with a
standard deviation of 807 seconds.

* The distributions of finishing times for both groups are approximately Normal.


** Remember: a better performance corresponds to a faster finish. **


### (a) Write down the short-hand for these two normal distributions.


For men:
$$ N(\mu =4313, \sigma = 538) $$
For women:
$$ N(\mu =5261, \sigma = 807) $$

###(b) What are the Z-scores for Leo’s and Mary’s finishing times? What do these Z-scores tell you?

We have a few options. The Z score can be calculated by:
$$ z = \frac{x-\bar{x}}{\sigma(X)} $$ 


```{r}
leoTime <- 4948
menAvg <- 4313
menStd <- 583 # Phrasing!
maryTime <- 5513
womenAvg <- 5261
womenStd <- 807 
## Just the variable
```
Now for the first calculation:
```{r}
leoZ <- (leoTime - menAvg)/menStd
maryZ <- (maryTime - womenAvg)/womenStd

if (leoZ == maryZ){
  print("They are, when normalized, equally as fast")
} else if (leoZ > maryZ){
  print("Leo is slower once normalized.")
} else 
  print("Mary is slower once normalized.")
```


###(c) Did Leo or Mary rank better in their respective groups? Explain your reasoning.

Because the rankings are normally distrobuted Mary did better as a percentage rank than Leo, but since we don't know the number of participants, Leo could have come in 10th and Mary could have come in 100th depending on the number of competitors. Assuming similar number of competitors, Mary would have a higher ranking. 


###(d)  What percent of the triathletes did Leo finish faster than in his group?


```{r}
pnorm(leoZ, lower.tail = FALSE) * 100 # Necessary to get the P[X > x].
```


### (e) What percent of the triathletes did Mary finish faster than in her group?

```{r}
pnorm(maryZ, lower.tail = FALSE) * 100 # Necessary to get the P[X > x].
```

### (f) If the distributions of finishing times are not nearly normal, would your answers to parts (b) - (e) change? Explain your reasoning.

The answers to part b) would not change because z scores can be calculated even without normalicy. Part c) we might be able to tell who ranked better, given the same number or participants, because if the women had a long right left tail and the men a long right tail, we would know that the women Mary was higher in percentage. We can not say anything WRT d)-f).

I'm a bit unsure of this last part because of the CLT. On slide 10 of [this](http://web.as.uky.edu/statistics/users/pbreheny/580-f10/notes/9.pdf) pdf, Breheny seems to say that percentiles could be aproximated. I am going to stick with what the book said though. 

Here is a really interesting link showing how non normal data can look like: https://www.autodeskresearch.com/publications/samestats




